Who does big team science?

### 2023-03-10 manuscript_scopus.Rmd
adding all the outputs

```{r data, include = FALSE}
geo_country <- import("py_outputs/geo_country.xlsx")
```

@@

-   Number of articles for inclusion: total, separated by four subject
    areas, presenting graphics of the number of publication across time
-   Number of distinct journals within each of the four subject areas
-   Statistics (mean, median, standard deviation, minimum, maximum) of
    the journal impact using Source Normalized Impact per Paper for the
    four subject areas.

@@

-   The totals of the number of articles published within the smaller
    subject area classifications. We will visualize these differences
    over time to determine if there is increasing interest in a specific
    subarea over time for each of the four larger subject area
    classifications.
-   The keywords present in the publications data overall to identify
    trends and common themes in the publications for the four subject
    areas using visualizations (wordclouds) to depict the common
    keywords.

@@

-   The total number of unique authors
-   Statistics (mean, standard deviation, minimum, maximum, median) on
    the number of authors included on publications.
-   We will present visualizations of these results across time.

@@

-   We will create a visualization of the trend and variance of
    researcher career length across publication years.
-   To analyze trends over time, we will calculate the average career
    length for each publication (i.e., average the author career length
    to create one score for each paper) and run a regression analysis
    using career length to predict year of publication. A positive slope
    for year of publication would indicate increasing years of first
    publication (i.e., more younger scholars over time), while a
    negative slope would indicate older years of first publication
    (i.e., more older scholars over time).
-   In order to show variance between individuals, we calculate the
    standard deviation of career length for each publication and run a
    regression analysis using this variance representation to predict
    publication year. A positive slope would indicate increasing
    variance over time (i.e., more diversity in the career lengths of
    scholars), while a negative slope would indicate less variance and
    diversity in scholars over time.
-   These analyses will be completed separately for each of the four
    large subject areas.

@@

@@

-   We will summarize the general education categories of individuals at
    the time of publication, along with a summary for change over time.

@@

@@

-   We will report descriptive statistics on the total number of
    publications, i10 index, and h-index for individuals overall.
-   Next, we will use the same analyses described in the career length
    section to analyze trends over time. An increasing slope over time
    indicates that individuals who are publishing more are more
    represented in BTS over time (i.e., increasing numbers of scholars
    with higher publication rates), while a negative slope indicates
    more researchers with less publications.
-   A positive slope for standard deviation indicates increasing
    variance over time (i.e., more diversity in the individual
    publication rates), while a negative slope would indicate less
    diversity in researchers over time. While publication rates do not
    represent value as a researcher, they are often used in hiring and
    promotion decisions, and we will use this variable as a proxy to
    gauge the diversity in scholars represented in big teams.

@@

-   Last, we will examine the differences in representation for
    corresponding author sets versus all other authors. For papers with
    10 to 49 authors, we will use the three first authors and the last
    author to compare against other authors. For 50 to 99 authors, five
    first authors plus last will be used, and for all papers with more
    than 100 authors, we will use ten first authors and the last author.
    We will calculate the frequencies of each of the UN Sub-Regions for
    first authors versus other authors, converting these values to
    proportions. Given the expected small sample sizes of these
    contingency tables, we will group together titles based on the year
    of publication (assuming at least 5 publications per year, these may
    be binned by 5-year or smaller increments to increase sample size).
    For each grouping, we will calculate the effect size of the
    differences in frequencies comparing first authors to all other
    authors. Since this data is categorical, we will use Cramer's V to
    represent the effect size. If the effect size includes zero in its
    confidence interval, this result will imply that first and all other
    authors represent the same pattern of UN Sub-Region diversity. Any
    confidence interval that does include zero represents a difference
    in diversity. We will report these values and discuss what regions
    of the world are represented when effect sizes indicate a different
    from zero using standardized residuals.

>>>>

```{r data, include = FALSE}
geo_country <- import("py_outputs/geo_country.xlsx")
geo_region <- import("py_outputs/geo_region.xlsx")
```

@@

-   Number of articles for inclusion: total, separated by four subject
    areas, presenting graphics of the number of publication across time

total

13581

+----------------+-----+
|publication_type|count|
+----------------+-----+
|              sh|   41|
|              cp|  266|
|              ar|12689|
|              dp|    4|
|              pp|  580|
|              ip|    1|
+----------------+-----+

subject area 
+-----------------+-----+
|      SubjectArea|count|
+-----------------+-----+
|  Health Sciences|10685|
|Physical Sciences| 6151|
|Multidisciplinary| 1569|
|  Social Sciences|  543|
|    Life Sciences| 7686|
+-----------------+-----+

time

pub_time_count

-   Number of distinct journals within each of the four subject areas

Out[52]: 2534

+-----------------+------------+
|      SubjectArea|count(srcid)|
+-----------------+------------+
|  Health Sciences|        1472|
|Physical Sciences|         641|
|Multidisciplinary|          14|
|  Social Sciences|         206|
|    Life Sciences|        1024|
+-----------------+------------+

-   Statistics (mean, median, standard deviation, minimum, maximum) of
    the journal impact using Source Normalized Impact per Paper for the
    four subject areas.

+-----------------+------------------+------------------+---------+------------+---------+
|      SubjectArea|         avg(snip)| stddev_samp(snip)|min(snip)|median(snip)|max(snip)|
+-----------------+------------------+------------------+---------+------------+---------+
|  Health Sciences|  4.39848449724717|7.1846479513521615|      0.0|       2.078|  173.932|
|Physical Sciences|1.7298434289762685| 1.336672368345959|    0.155|       1.278|   30.405|
|Multidisciplinary| 6.092944123314063|   4.3151334953509|    0.032|       3.341|   11.342|
|  Social Sciences| 2.204784860557769|2.1441319264666743|    0.017|       1.605|   30.405|
|    Life Sciences| 2.524247249015079| 2.044898988084329|    0.017|       1.742|   12.624|
+-----------------+------------------+------------------+---------+------------+---------+

@@

-   The totals of the number of articles published within the smaller
    subject area classifications. We will visualize these differences
    over time to determine if there is increasing interest in a specific
    subarea over time for each of the four larger subject area
    classifications.

pub_time_subject

-   The keywords present in the publications data overall to identify
    trends and common themes in the publications for the four subject
    areas using visualizations (wordclouds) to depict the common
    keywords.

keywords

@@

-   The total number of unique authors

+-----------+
|count(auid)|
+-----------+
|     257536|
+-----------+

-   Statistics (mean, standard deviation, minimum, maximum, median) on
    the number of authors included on publications.

	count(Authorseq)	avg(Authorseq)	stddev_samp(Authorseq)	min(Authorseq)	median(Authorseq)	max(Authorseq)
0	13581	49.904646	207.006114	10	19.0	5155

-   We will present visualizations of these results across time.

author_time

@@

-   We will create a visualization of the trend and variance of
    researcher career length across publication years.

career_time

-   To analyze trends over time, we will calculate the average career
    length for each publication (i.e., average the author career length
    to create one score for each paper) and run a regression analysis
    using career length to predict year of publication. A positive slope
    for year of publication would indicate increasing years of first
    publication (i.e., more younger scholars over time), while a
    negative slope would indicate older years of first publication
    (i.e., more older scholars over time).
-   In order to show variance between individuals, we calculate the
    standard deviation of career length for each publication and run a
    regression analysis using this variance representation to predict
    publication year. A positive slope would indicate increasing
    variance over time (i.e., more diversity in the career lengths of
    scholars), while a negative slope would indicate less variance and
    diversity in scholars over time.
-   These analyses will be completed separately for each of the four
    large subject areas.

------ Physical Sciences ------

Note: the last rows are the information for Intercept
## -------------------------------------------------
##   Estimate   |   Std.Error | t Values  |  P-value
##  -0.531373   0.030990  -17.146   0.000000
##  -0.207000   0.059698   -3.467   0.000533
## 2020.342936   0.367600 5496.040   0.000000
## ---
## Mean squared error:  30.862113 , RMSE:  5.555368
## Multiple R-squared: 0.145118 ,             Total iterations: 0
------------

------ Social Sciences ------

Note: the last rows are the information for Intercept
## -------------------------------------------------
##   Estimate   |   Std.Error | t Values  |  P-value
##  -0.437422   0.093563   -4.675   0.000004
##  -0.104549   0.146459   -0.714   0.475750
## 2019.182625   0.764850 2639.973   0.000000
## ---
## Mean squared error:  30.685713 , RMSE:  5.539469
## Multiple R-squared: 0.118772 ,             Total iterations: 0
------------

------ Health Sciences ------

Note: the last rows are the information for Intercept
## -------------------------------------------------
##   Estimate   |   Std.Error | t Values  |  P-value
##  -0.460876   0.019674  -23.425   0.000000
##  -0.218675   0.037611   -5.814   0.000000
## 2019.076458   0.231406 8725.260   0.000000
## ---
## Mean squared error:  30.039405 , RMSE:  5.480822
## Multiple R-squared: 0.144584 ,             Total iterations: 0
------------

------ Life Sciences ------

Note: the last rows are the information for Intercept
## -------------------------------------------------
##   Estimate   |   Std.Error | t Values  |  P-value
##  -0.604957   0.026519  -22.812   0.000000
##  -0.056843   0.046915   -1.212   0.225716
## 2019.490201   0.276602 7301.067   0.000000
## ---
## Mean squared error:  27.654316 , RMSE:  5.258737
## Multiple R-squared: 0.175747 ,             Total iterations: 0
------------

@@

+----------------------+
|count(affiliation_ids)|
+----------------------+
|                 35114|
+----------------------+

inst time

@@

We will summarize the general education categories of individuals at the time of publication, along with a summary for change over time.

    Figure out how messy this is for text processing requirements
    For each publication, plot of proportion of doctorates across time - if the number is going down, we are involving more diverse types of people (assuming that doctorates are "normal" academics/researchers)
    We will only perform this analysis if the proportion of information avaliable is over 50%.

+-----------+------+
|    degrees| count|
+-----------+------+
|       null|656440|
|         MD|  8190|
|        PhD|  3879|
|    MD, PhD|  1267|
|       M.D.|  1002|
|        Dr.|   490|
|        MSc|   461|
|        MPH|   368|
| Prof., PhD|   349|
|      Ph.D.|   348|
|         MS|   330|
|     MD PhD|   285|
|        MD.|   279|
|   Prof. MD|   266|
|  Prof. PhD|   239|
|         BS|   232|
|       Prof|   220|
|         BA|   196|
|M.D., Ph.D.|   191|
|    MD, MPH|   158|
+-----------+------+

@@

pub_type versus normal_pub_type

@@

    We will report descriptive statistics on the total number of publications and h-index for individuals overall.
        Do this for each unique person and report averages
        Do this for each publication and create an average for each publication
    Next, we will use the same analyses described in the career length section to analyze trends over time. An increasing slope over time indicates that individuals who are publishing more are more represented in BTS over time (i.e., increasing numbers of scholars with higher publication rates), while a negative slope indicates more researchers with less publications.
    A positive slope for standard deviation indicates increasing variance over time (i.e., more diversity in the individual publication rates), while a negative slope would indicate less diversity in researchers over time. While publication rates do not represent value as a researcher, they are often used in hiring and promotion decisions, and we will use this variable as a proxy to gauge the diversity in scholars represented in big teams.
    We will separate these by each of the four subject areas.

	avg_pub_count	std_pub_count
0	4.451816	6.901369

	avg(avg_pub_count AS mean_of_avg_pub_count)	stddev_samp(avg_pub_count AS std_of_avg_pub_count)	avg(std_pub_count AS mean_of_std_pub_count)	stddev_samp(std_pub_count AS std_of_std_pub_count)
0	7.581413	5.797233	7.140015	5.10383

	avg_HIndex	std_HIndex	med_HIndex
0	14.242142	43.798203	2.0

	avg_of_avg_HIndex	std_of_avg_HIndex	avg_of_std_HIndex	std_of_std_HIndex	med_of_med_HIndex
0	28.136255	38.013819	36.430509	46.474619	5.0

------ Physical Sciences ------

Note: the last rows are the information for Intercept
## -------------------------------------------------
##   Estimate   |   Std.Error | t Values  |  P-value
##  -0.004420   0.022715   -0.195   0.845717
##  -0.105746   0.034058   -3.105   0.001922
## 2015.347578   0.195556 10305.717   0.000000
## ---
## Mean squared error:  35.756076 , RMSE:  5.979638
## Multiple R-squared: 0.009554 ,             Total iterations: 0
------------

------ Social Sciences ------

Note: the last rows are the information for Intercept
## -------------------------------------------------
##   Estimate   |   Std.Error | t Values  |  P-value
##  -0.335536   0.179536   -1.869   0.062389
##   0.130796   0.186924    0.700   0.484517
## 2016.658826   0.487148 4139.727   0.000000
## ---
## Mean squared error:  34.070868 , RMSE:  5.837026
## Multiple R-squared: 0.021558 ,             Total iterations: 0
------------

------ Health Sciences ------

Note: the last rows are the information for Intercept
## -------------------------------------------------
##   Estimate   |   Std.Error | t Values  |  P-value
##  -0.235551   0.028158   -8.365   0.000000
##  -0.006692   0.026612   -0.251   0.801476
## 2014.979175   0.119976 16794.893   0.000000
## ---
## Mean squared error:  33.860172 , RMSE:  5.818949
## Multiple R-squared: 0.035782 ,             Total iterations: 0
------------

------ Life Sciences ------

Note: the last rows are the information for Intercept
## -------------------------------------------------
##   Estimate   |   Std.Error | t Values  |  P-value
##  -0.430599   0.037807  -11.389   0.000000
##   0.133215   0.034255    3.889   0.000102
## 2016.008237   0.147870 13633.624   0.000000
## ---
## Mean squared error:  31.736306 , RMSE:  5.633499
## Multiple R-squared: 0.054081 ,             Total iterations: 0
------------

@@

total number of unique GP for each publication 
using year to predict GP 
positive slope means that as year goes up number of GP goes up 
negative slope means that as year goes up number of GP goes down 
forest plot of the effects + confidence interval

------ Physical Sciences ------

Note: the last rows are the information for Intercept
## -------------------------------------------------
##   Estimate   |   Std.Error | t Values  |  P-value
##  -0.061001   0.022958   -2.657   0.007923
## 130.518393  46.247823    2.822   0.004802
## ---
## Mean squared error:  56.853041 , RMSE:  7.540096
## Multiple R-squared: 0.002357 ,             Total iterations: 0
------------

------ Social Sciences ------

Note: the last rows are the information for Intercept
## -------------------------------------------------
##   Estimate   |   Std.Error | t Values  |  P-value
##  -0.077040   0.058153   -1.325   0.186023
## 160.766462 117.214988    1.372   0.170993
## ---
## Mean squared error:  45.807761 , RMSE:  6.768143
## Multiple R-squared: 0.004491 ,             Total iterations: 0
------------

------ Health Sciences ------

Note: the last rows are the information for Intercept
## -------------------------------------------------
##   Estimate   |   Std.Error | t Values  |  P-value
##  -0.017775   0.009794   -1.815   0.069578
##  40.118948  19.718323    2.035   0.041926
## ---
## Mean squared error:  25.532574 , RMSE:  5.052977
## Multiple R-squared: 0.000434 ,             Total iterations: 0
------------

------ Life Sciences ------

Note: the last rows are the information for Intercept
## -------------------------------------------------
##   Estimate   |   Std.Error | t Values  |  P-value
##  -0.014946   0.010485   -1.425   0.154081
##  34.499812  21.117788    1.634   0.102392
## ---
## Mean squared error:  17.279398 , RMSE:  4.156850
## Multiple R-squared: 0.000434 ,             Total iterations: 0
------------

-   Last, we will examine the differences in representation for
    corresponding author sets versus all other authors. For papers with
    10 to 49 authors, we will use the three first authors and the last
    author to compare against other authors. For 50 to 99 authors, five
    first authors plus last will be used, and for all papers with more
    than 100 authors, we will use ten first authors and the last author.
    We will calculate the frequencies of each of the UN Sub-Regions for
    first authors versus other authors, converting these values to
    proportions. Given the expected small sample sizes of these
    contingency tables, we will group together titles based on the year
    of publication (assuming at least 5 publications per year, these may
    be binned by 5-year or smaller increments to increase sample size).
    For each grouping, we will calculate the effect size of the
    differences in frequencies comparing first authors to all other
    authors. Since this data is categorical, we will use Cramer's V to
    represent the effect size. If the effect size includes zero in its
    confidence interval, this result will imply that first and all other
    authors represent the same pattern of UN Sub-Region diversity. Any
    confidence interval that does include zero represents a difference
    in diversity. We will report these values and discuss what regions
    of the world are represented when effect sizes indicate a difference
    from zero using standardized residuals.

```{r}
library(tidyr)
geo_region_wide <- geo_region %>% 
  pivot_wider(id_cols = Year, 
              names_from = aff_un, 
              values_from = count,
              values_fill = 0)
```

end up with a list of effect sizes by year by small medium large 
x axis is year
y axis is the effect size and the confidence interval 
if diversity is increasing then the patterns would be more similar (effect size of zero)
if diversity is not increasing then the patterns would be different (effect size does not include zero)

who is not represented in the "corresponding" spots - proportion chart by decades stacked bar chart of the decades
